---
title: "dataframe"
author: "Nicholus Tint Zaw"
date: "2022-08-14"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# What we are going to cover in this exercise? 

1. Creating a Data Frame
1. Exploring dataframe
1. Dataframe indexing 
1. Running Functions on Dataframes



# Creating a Data Frame 

## as.dataframe 

```{r}
# Definition of vectors
# ref: https://towardsdatascience.com/introduction-to-data-frames-in-r-b9a6302d9a56

name <- c("Mercury", "Venus", "Earth", "Mars", 
          "Jupiter", "Saturn", "Uranus", "Neptune")

type <- c("Terrestrial planet", "Terrestrial planet", 
          "Terrestrial planet", "Terrestrial planet", 
          "Gas giant", "Gas giant", "Gas giant", "Gas giant")

diameter <- c(0.382, 0.949, 1, 0.532, 11.209, 9.449, 4.007, 3.883)

rotation <- c(58.64, -243.02, 1, 1.03, 0.41, 0.43, -0.72, 0.67)

rings <- c(FALSE, FALSE, FALSE, FALSE, TRUE, TRUE, TRUE, TRUE)

# Create a data frame from the vectors
planets_df <- data.frame(name, type, diameter, rotation, rings)

```



# Exploring dataframe 

```{r}
# import base-dataframe mtcars
df <- mtcars 

# view dataframe
View(planets_df) 

View(df)

# return the column names
names(planets_df)

names(df)

# print subset of dataframe
head(df)

tail(df)

# dimension of dataframe
dim(df) # returns the number of rows first, then the number of columns.

# number of observation in dataframe
length(df$mpg)

# check below syntax 
length(df)


# class of dataframe and columns
class(df)

class(df$cyl)

# check structure of dataframe 
str(df) 
str(planets_df)

```



# Dataframe indexing

## using square brackets [] 

```{r}
# Return the value in the first row and first column:
df[1,1]

# Return the value in the second row and first column:
df[2,1]

# Return the value in the third row and second column:
df[3,2]

# Return all the values in the first row:
df[1,]

# Return the values in the first through seventh rows, in the second column:
mtcars[1:7,2]

```

## use the $ operator to refer to the column within the dataframe 

```{r}
df$cyl

df$wt
```

## Some advance indexing 

application of condition in indexing 

```{r}
# return only the rows of data with cars that have four cylinders
df[which(df$cyl == 4),]

# return the cars with more than 90 horsepower.
df[which(df$hp > 90),]

# return the cars with high mpg and low weight.
df[which(df$mpg > 28 & df$wt < 2),]

# use indexing on just one column of a dataframe 
df$mpg[which(df$cyl == 4)]

# run functions on subsets of a column of data
mean(df$mpg[which(df$cyl == 4)])

```




# Running Functions on Dataframes

```{r}
# How many values are in the mpg column?
length(df$mpg)

# What is the average horsepower of these cars?
mean(df$hp)

# What is the range of the weights of these cars?
range(df$wt)

# What is the frequency of cylinder type?
table(df$cyl)

table(planets_df$type)
```

