---
title: "recode variables"
author: "Nicholus Tint Zaw"
date: "2022-08-20"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# What we are going to cover in this exercise? 

1. ifelse()
1. cut()
1. case_when()

# setting dataframe 

```{r}
df <- iris 
names(iris)

summary(df$Sepal.Length)
min(df$Sepal.Length)
max(df$Sepal.Length)
mean(df$Sepal.Length)
median(df$Sepal.Length)

quantile(df$Sepal.Length)
quantile(df$Sepal.Length, probs = 0.25) # 1st quartile 
quantile(df$Sepal.Length, probs = 0.50) # 2nd quartile 
quantile(df$Sepal.Length, probs = 0.75) # 3rd quartile 

```


# ifelse() 

```{r}
# with string variable 
df$versicolor <- ifelse(df$Species == "versicolor", "versicolor", "other")

# with numeric variable (>= median = long, < median = short)
df$Sepal.Length.Dummy <- ifelse(df$Sepal.Length >= median(df$Sepal.Length), 
                                "long", "short")

# nested ifelse
# (assigned Sepal.Length into three different quantile category)
q1 <- quantile(df$Sepal.Length, probs = 0.25)
q2 <- quantile(df$Sepal.Length, probs = 0.50)
q3 <- quantile(df$Sepal.Length, probs = 0.75)

df$Sepal.Length.Q <- ifelse(df$Sepal.Length < q1, 
                            "1st quarter", 
                            ifelse(df$Sepal.Length >= q1 & df$Sepal.Length < q2, 
                                   "2nd quarter",
                                   ifelse(df$Sepal.Length >= q2 & df$Sepal.Length < q3,
                                          "3rd quarter", 
                                          ifelse(df$Sepal.Length >= q3, 
                                                 "4th quarter", NA)
                                          )))

table(df$Sepal.Length.Q)
```


# cut() 

```{r}
x <- 1:21

quantile(x)

```
If single vector value was specified at `breaks` parameter, it will equally divided the interested vectors into the number of category specified at `breaks` parameter. 

```{r}
cut(x, 3)
cut(x, 3, dig.lab = 4, ordered = TRUE) # order the levels 

```


We can specify the interval criteria by using vector in `breaks` parameter. 

```{r}
# check the result! is that what we want? 
cut(x, breaks = c(7, 14, 21))

# revised the syntax a little bit and check the result!
cut(x, breaks = c(1, 7, 14, 21))

cut(x, breaks = c(0, 7, 14, 21))
cut(x, breaks = c(1, 7, 14, 21), include.lowest = TRUE)

```


```{r}
quantile(x)

cut(x, breaks = c(1, 6, 11, 16, 21), include.lowest = TRUE)

cut(x, 
    breaks = c(1, 6, 11, 16, 21), 
    include.lowest = TRUE, 
    labels = c("1st Q", "2nd Q", "3rd Q", "4th Q"))

```

Check the result, when we applied the `right` option parameter. Is it still what we want? 

```{r}

cut(x, 
    breaks = c(1, 6, 11, 16, 21), 
    include.lowest = TRUE)

cut(x, 
    breaks = c(1, 6, 11, 16, 21), 
    include.lowest = TRUE, 
    right = FALSE)
```

If we are not interest about the higher value in the distribution and don't want to specify in the `breaks` parameter vector; 

```{r}
x <- 1:100

# only assigned pass/fail using 40 as cut-off point

cut(x, 
    breaks = c(0, 40, Inf), 
    include.lowest = TRUE, 
    labels = c("fail", "pass"))


```



# case_when()

We are going to use `case_when()` function from `dplyr` package. 

```{r, error = FALSE}
library(dplyr)
```

```{r}
x <- 1:50

y <- 
  case_when(
  x < 5  ~ "< 5",
  x == 7  ~ "equal to 7",
  x > 35  ~ "> 35"
)


```


```{r}
# check the result and interpret the result! 

case_when(
  x %% 35 == 0 ~ "thridy five",
  x %% 5 == 0 ~ "five",
  x %% 7 == 0 ~ "seven",
  TRUE ~ as.character(x)
)

# what is %%? 
# ref: https://stackoverflow.com/questions/30257819/what-does-the-double-percentage-sign-mean
```

