---
title: "dataframe"
author: "Nicholus Tint Zaw"
date: "2022-08-14"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# What we are going to cover in this exercise? 

1. Function


```{r}
set.seed(523535)

x <- rnorm(100)

# apply based functions
median(x)
mean(x)
min(x)
max(x)
quantile(x, 0.25)
quantile(x, c(0.25, 0.75))
summary(x)

```

Now, it is time to create your own function!


$\bar{X} = \frac{\sum x_i}{N}$

What input parameter do we need to calculate mean? Of course, the numeric vector, it length (number of observation in the vector), and summation of all numbers from that vector. Right? 

```{r}

my_mean <- function(x){ # x = input parameter - numeric vector 
  
  # n = number of observation in vector x 
  n <- length(x)
  
  # summation of vector 
  sumvect <- sum(x)
  
  # mean calculation 
  cal_mean <- round(sumvect/n, 3)
  
  return(cal_mean)
  
}

```


Now, please working yourself on SD function!

$\sigma = \sqrt{\frac{\sum (x_i - \bar{X})^2}{N - 1}}$


```{r}

my_sd <- function(x){
  
  # n = number of observation in vector x 
  n <- length(x)
  
  # summation of vector 
  sumvect <- sum(x)
  
  # mean calculation 
  cal_mean <- round(sumvect/n, 3)
  
  # variance calculation 
  cal_var <- sum((x - cal_mean)^2)/(n-1)
  
  # sd calculation 
  cal_sd = sqrt(cal_var)
  
  return(cal_sd)
  
}

```



Then, create a function which return the `dataframe` with contain `category variable` as index, `frequency` and `proportion` as column for each respective category. Please use `mtcars` data as sample dataset. 

```{r}
df <- mtcars

table(df$cyl)

prop.table(table(df$cyl))


my_table <- function(input){
  
  frequency <- table(input)

  proportion <- prop.table(table(input)) * 100
  
  rpt_tbl <- cbind(frequency, proportion)
  
  return(rpt_tbl)

}

my_table(df$cyl)

my_table(df$mpg)

```


```{r}

freqency <- table(df$cyl)
proportion <- prop.table(table(df$cyl))
result_table <- cbind(freqency, proportion)


```



## crate table as dataframe (quantile category)

```{r}
my_qrt_table <- function(input){
  
  # calculate quantile values 
  q1 <- quantile(input, probs = 0.25)
  q2 <- quantile(input, probs = 0.50)
  q3 <- quantile(input, probs = 0.75)

  # crate category variable 
  qrt_cat <- ifelse(input < q1, "1st quarter", 
                    ifelse(input >= q1 & input < q2, "2nd quarter", 
                           ifelse(input >= q2 & input < q3, "3rd quarter", 
                                  ifelse(input >= q3, "4th quarter", NA))))
  
  frequency <- table(qrt_cat)

  proportion <- prop.table(table(qrt_cat)) * 100
  
  rpt_tbl <- cbind(frequency, proportion)
  
  return(rpt_tbl)

}

my_qrt_table(iris$Petal.Width)
my_qrt_table(mtcars$mpg)


```

