---
title: "looping"
author: "Nicholus Tint Zaw"
date: "2022-08-31"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```



# For loops

ref: https://r4ds.had.co.nz/iteration.html#iteration

Let's say we wnat to calculate the mediam value for each vector from the tibble. 

```{r}
# create tilbble

df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

median(df$a)

median(df$b)

median(df$c)

median(df$d)

```

Let apply loop instead of individual command for each vector. 

```{r}

output <- vector("double", ncol(df))  # 1. output

for (i in seq_along(df)) {            # 2. sequence
  output[[i]] <- median(df[[i]])      # 3. body
}

output

```

## Let's do some exercise!

1. Compute the mean of every column in `mtcars`. 
1. Determine the type of each column in `nycflights13::flights`.
1. Compute the number of unique values in each column of `iris`.


```{r}
# data type ref: https://www.postgresql.org/docs/current/datatype-numeric.html 

# Compute the mean of every column in `mtcars`. 
mtcars_means <- vector("double", ncol(mtcars))  # 1. output

for (i in seq_along(mtcars)) {            # 2. sequence
  mtcars_means[[i]] <- mean(mtcars[[i]])        # 3. body
}

mtcars_means


# Determine the type of each column in `nycflights13::flights`.
df <- nycflights13::flights

flight_type <- vector("double", ncol(df))  # 1. output

for (i in seq_along(df)) {            # 2. sequence
  
  flight_type[[i]] <- typeof(df[[i]])        # 3. body
}

flight_type


# Compute the number of unique values in each column of `iris`.
iris_unique <- vector("double", ncol(iris))  # 1. output

for (i in seq_along(iris)) {            # 2. sequence
  iris_unique[[i]] <- length(unique(iris[[i]]))        # 3. body
}

iris_unique

```



## Unknown output length 

```{r}
means <- c(0, 1, 2)

output <- double()

for (i in seq_along(means)) {
  n <- sample(100, 1)
  output <- c(output, rnorm(n, means[[i]]))
}

str(output)


```

```{r}
out <- vector("list", length(means))

for (i in seq_along(means)) {
  n <- sample(100, 1)
  out[[i]] <- rnorm(n, means[[i]])
}

str(out)


str(unlist(out))

```

## Unknown sequence length 

```{r}

flip <- function() sample(c("T", "H"), 1)

flips <- 0
nheads <- 0

while (nheads < 3) {
  
  if (flip() == "H") {
    nheads <- nheads + 1
  } else {
    nheads <- 0
  }
  
  flips <- flips + 1
}

flips



```


## HW 

dice - roll 

looping will run till we got 6, 10 time

pls report - number of rolls, mean value of n rolls results 


```{r}
set.seed(34534)

rolldice <- function() sample(c(1:52), 1)

output <- double()

nroll <- 0
nsix <- 0

while (nsix < 5) {
  
  result <- rolldice()
  
  if (result == 21) {
    nsix <- nsix + 1
  } else {
    nsix <- nsix
  }
  
  nroll <- nroll + 1
  output <- c(output, result)
}


nroll
summary(output)

```



## any shortcuts? 

Of course! yes. 

```{r}
df %>% map_dbl(mean)

df %>% map_dbl(median)

df %>% map_dbl(sd)

```
Other function from `map` functions from `purrr` package.  

1. map() makes a list.
1. map_lgl() makes a logical vector.
1. map_int() makes an integer vector.
1. map_dbl() makes a double vector.
1. map_chr() makes a character vector.

```{r}

models <- mtcars %>% 
  split(.$cyl) %>% 
  map(function(df) lm(mpg ~ wt, data = df))


testm <- iris %>%
  split(.$Species) %>%
  map(function(df) mean(df$Sepal.Length))

unique(mtcars$cyl)

summary(models[[2]])

cyl_6 <- mtcars %>%
  filter(cyl == 6) 

model1 <- lm(mpg ~ wt, cyl_6)
summary(model1)

```



```{r}
models <- mtcars %>% 
  split(.$cyl) %>% 
  map(~lm(mpg ~ wt, data = .))

testm2 <- iris %>% 
  split(.$Species) %>% 
  map(~mean(.$Sepal.Length))


models %>% 
  map(summary) %>% 
  map_dbl(~.$r.squared)

```


## Base R function 

### apply 

```{r}

# create sample data
sample_matrix <- matrix(C<-(1:10),nrow=3, ncol=10)
  
print( "sample matrix:")
sample_matrix
  
# Use apply() function across row to find sum
print("sum across rows:")
apply(sample_matrix, 1, sum)
  
# use apply() function across column to find mean
print("mean across columns:")
apply( sample_matrix, 2, mean)

```

### lapply 

```{r}
# create sample data
names <- c("priyank", "abhiraj","pawananjani", "sudhanshu","devraj")
print( "original data:")
names

# apply lapply() function
print("data after lapply():")
lapply(names, toupper)


x <- list(a = 1:10, beta = exp(-3:3), logic = c(TRUE,FALSE,FALSE,TRUE))
# compute the list mean for each list element
lapply(x, mean)


lapply(out, mean)


```

### sapply

```{r}
 
x1 <- list(
  c(0.27, 0.37, 0.57, 0.91, 0.20),
  c(0.90, 0.94, 0.66, 0.63, 0.06), 
  c(0.21, 0.18, 0.69, 0.38, 0.77)
)
x2 <- list(
  c(0.50, 0.72, 0.99, 0.38, 0.78), 
  c(0.93, 0.21, 0.65, 0.13, 0.27), 
  c(0.39, 0.01, 0.38, 0.87, 0.34)
)

threshold <- function(x, cutoff = 0.8) x[x > cutoff]

x1 %>% sapply(threshold) %>% str()

x2 %>% sapply(threshold) %>% str()

```


### tapply 

```{r}

# print head of diamonds dataset
print(" Head of data:")
head(diamonds)


unique(diamonds$cut)

# apply tapply function to get average price by cut
print("Average price for each cut of diamond:")
tapply(diamonds$price, diamonds$cut, mean)


tapply(iris$Sepal.Length, iris$Species, mean)

tapply(iris$Sepal.Length, iris$Species, summary)

tapply(mtcars$mpg, mtcars$cyl, median)


```

